Associations between chlorophyll \emph{a} and various microcystin health advisory Concentrations

### 2016-06-07 manuscript/manuscript.tex
updates, including proofs

Over the last decade, numerous events and legislative activities have
raised the public awareness of harmful algal blooms {[}1--3{]}. In
response the US Environmental Protection Agency (USEPA) has recently
released suggested microcystin (one of the more common toxins)
concentrations that would trigger health advisories {[}4--6{]}.
Additionally, the World Health Organization (WHO) has proposed
microcystin advisory levels for drinking water and a range of
recreational risk levels {[}7,8{]}. While these levels and associated
advisories are likely to help mitigate the impacts from harmful algal
blooms, they are not without complications.

@@

Fortunately, cyanobacteria and microcystin-LR has been shown to be
associated with several other, more commonly measured and well
understood components of water quality that are readily assessed in the
field {[}11{]}. For instance, there are small or hand held fluorometers
that measure chlorohpyll \emph{a}. Additionally, chlorophyll \emph{a} is
a very commonly measured component of water quality that is also known
to be positively associated with microsystin-LR concentrations
{[}12,13{]}. Recently, Yuan et. al {[}13{]} explored these associations
in detail and controlled for other related variables. In their analysis
they find that total nitrogen and chlorophyll \emph{a} show the
strongest association with microcystin. Furthermore, they identify
chlorophyll\emph{a} and total nitrogen concentrations that are
associated with exceeding 1 \(\mu\)g/L of microcystin. These findings
suggest that chlorophyll a concentrations could also track the new USEPA
microcystin health advisory levels for drinking water. Identifying this
association would provide an important tool for water resource managers
to help manage the threat to public health posed by cHABs and would be
especially useful in the absence of measured microcystin concentrations.

In this paper we build on past efforts and utilize the National Lakes
Assessment (NLA) data and identify chlorophyll \emph{a} concentrations
that are associated with higher probabilities of exceeding several
microcystin health advisory concentrations {[}6,8,14{]}. We build on
past studies by exploring associations with the newly announced advisory
levels and by also applying a different method, conditional probability
analysis. Utilizing different methods strengthens the evidence for
suggested chlorophyll \emph{a} levels that are associated with increased
risk of exceeding the health advisory levels as those levels are not
predicated on a single analytical method. So that others may repeat or
adjust this analysis, the data, code, and this manuscript are freely
available via
\href{https://github.com/USAPE/microcystinchla}{https://github.com/USEPA/microcystinchla}.

@@

We used the 2007 NLA chlorophyll \emph{a} and microcystin-LR
concentration data {[}14{]}. These data represent a snapshot of water
quality from the summer of 2007 for the conterminous United States and
were collected as part of an ongoing probabilistic monitoring program
{[}14{]}. Water quality data, including chlorophyll \emph{a} and
microcystin-LR were obtained via an intergated sample taken from the
surface of the lake down to 2 meters. Samples were taken at the same
time from the index site (e.g.~near the centroid of the lake) and these
provide the source for both chlorophyll \emph{a} and microcystin-LR
{[}14{]}.

@@

We used a conditional probability analysis (CPA) approach to explore
associations between chlorophyll \emph{a} concentrations and World
Health Organization (WHO) and USEPA microcystin health advisory levels
{[}16{]}. Many health advisory levels have been suggested (Table
\ref{tab:microcystin_levels}), but lakes with higher microcystin-LR
concentrations in the NLA were rare. Only 1.16\% of lakes sampled had a
concentration greater than 10 \(\mu\)g/L. Thus, for this analysis we
focused on the microcystin concentrations that are better represented in
the NLA data. These were the USEPA children's (i.e.~bottle fed infants
to pre-school age children) drinking water advisory level of 0.3
\(\mu\)g/L (USEPA Child), the WHO drinking water advisory level of 1
\(\mu\)g/L (WHO Drinking), the USEPA adult (i.e.~beyond pre-school aged
individuals) drinking water advisory level of 1.6 \(\mu\)g/L (USEPA
Adult), and the WHO recreational, low probability of effect advisory
level of 2 \(\mu\)g/L (WHO Recreational).

Conditional probability analysis provides information about the
probability of observing one event given another event has also occured.
For this analysis, we used CPA to examine how the conditional
probability of exceeding one of the health advisories changes as
chlorophyll \emph{a} increases in a lake. We expect to find higher
chlorohpyll \emph{a} concentrations to be associated with higher
probabilities of exceeding the microcystin health advisory levels. We
also calculated bootstrapped 95\% confidence intervals (CI) using 1000
bootstrapped samples. Thus, to identify chlorophyll \emph{a}
concentrations of concern we identified the value of the upper 95\% CI
across a range of conditional probabilities of exceeding each health
advisory level. Using the upper confidence limit to identify a threshold
is justified as it ensures that a given threshold is unlikely to miss a
microcystin exceedance.

As both microcystin-LR and chlorophyll \emph{a} values were highly right
skewed, a log base 10 transformation was used. Additional details of the
specific implementation are available at
\url{https://github.com/USEPA/microcystinchla}. A more detailed
discussion of CPA is beyond the scope of this paper, but see Paul et al.
{[}17{]} and Hollister et al. {[}18{]} for greater detail. Lastly, all
analyses were conducted using R version 3.2.2 and code and data from
this analysis are freely available as an R package at
\href{https://github.com/USAPE/microcystinchla}{https://github.com/USEPA/microcystinchla}.

@@

In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225
\(\mu\)g/L. Microcystin-LR concentrations of 0.05 \(\mu\)g/L represent
the detection limits. Any value greater than that indicates the presence
of microcystin-LR. Of those lakes with microcystin, the median
concentration was 0.51\(\mu\)g/L and the mean was 3.17\(\mu\)g/L. Of all
lakes sampled, 21\% of lakes exceeded the USEPA Child level, 8.8\% of
lakes exceeded the USEPA Adult level, 11.7\% of lakes exceeded the WHO
Drinking level, and 7.3\% of lakes exceeded the WHO Recreational level.
Chlorophyll \emph{a}, ranged from 0.07 to 936 \(\mu\)g/L and this
captures the range of trrophic states from oligotrophic to
hypereutrophic. All lakes had detectable levels of chlorophyll \emph{a}.
The median concentration was 7.79 \(\mu\)g/L and the mean was 29.63
\(\mu\)g/L. The association between chlorophyll \emph{a} and the upper
confidence interval across a range of conditional probability values are
shown in Table \ref{tab:mc_chla_table}. Specific chlorophyll \emph{a}
that are associated with greater than even odds of exceeding the
advisory levels were 23, 68, 84, and 104\(\mu\)g/L for 0.3, 1.0, 1.6,
and 2.0 \(\mu\)g/L advisory levels, respectively (Table
\ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).

The chlorophyll \emph{a} cutoffs may be used to predict whether or not a
lake exceeds the microcystin health advisories. Doing so allows us to
compare the accuracy of the prediction as well as evaluate false
negatives. Total accuracy of the four cutoffs predicting microcystin
exceedances were 74\% for the USEPA children's drinking water advisory,
86\% for the WHO drinking water advisory, 89\% for the USEPA adult
drinking water advisory, and 91\% for the WHO regreational advisory
(Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}).
However, total accuracy is only one part of the prediction performace
with which we are concerned.

@@

A second potential use is to identify past bloom events from historical
data. As harmful algal blooms are made up of many species and have
various mechanisms responsible for adverse impacts (e.g., toxins,
hypoxia, odors), there is no single definition of a bloom. For cHABs,
one approach has been to utilize phycocyanin to screen for or identify
bloom events {[}19{]}. This is a useful approach, but phycocyanin is not
always available, thus limiting its utility especially for examining
historical data. Using our chlorophyll \emph{a} cutoffs provides a value
that is also associated with microcystin-LR and can be used to classify
lakes, from past surveys, as having bloomed.

The values we propose are national and may miss regional variation in
water quality, including, chlorophyll \emph{a} and microcystin-LR
{[}21{]}. A set of regional conditional probabilities would be
interesting; however, limiting the analysis to the data available per
region would make interpretaion difficult. The sample size for each of
the conditional probabilities would be reduced (it ranges from 67 to
155) and the number of lakes in each region that exceed the microcystin
values is also reduced. The result is that our confidence in the
conditional probabilites would be less (i.e.~greatly increased
confidence intervals) and the relationships less pronounced as we have
fewer lakes on which to base the probabilites. Thus, this dataset is
best for making national scale recommendations.

Two other limitations with the 2007 National Lakes Assessment data are
that they represent a single sample from that lake and do not capture
temporal dynamics and without subsetting the data do not provide us the
ability to validate the presence of microcystin-LR. As of this writing,
the 2012 National Lakes Assessment data are not public. When these data
are released, a validation of this approach can be completed then.

Lastly, using chlorophyll \emph{a} is not meant as a replacement for
testing of microcystin-LR or other toxins. It should be used when other,
direct measurements of cyanotoxins are not available. In those cases,
which are likely to be common at least in the near future, using a more
ubiquitous measurement, such as chlorophyll \emph{a} will provide a
reasonable proxy for the probability of exceeding a microcystin health
advisory level and provide better protection against adverse effects in
both drinking and recreational use cases.

@@

\section{Figures}\label{figures}

\begin{verbatim}
## Saving 6 x 6.8 in image
\end{verbatim}

@@

\begin{longtable}[c]{@{}lll@{}}
\caption{Various suggested microcystin health advisory concentrations.
\label{tab:microcystin_levels}}\tabularnewline
\toprule
Source & Type & Concentration\tabularnewline
\midrule
\endfirsthead
\toprule
Source & Type & Concentration\tabularnewline
\midrule
\endhead
USEPA & Child Drinking Water Advisory & 0.3 \(\mu\)g/L\tabularnewline
WHO & Drinking Water & 1 \(\mu\)g/L\tabularnewline
USEPA & Adult Drinking Water Advisory & 1.6 \(\mu\)g/L\tabularnewline
WHO & Recreational: Low Prob. of Effect & 2-4 \(\mu\)g/L\tabularnewline
WHO & Recreational: Moderate Prob. of Effect & 10-20
\(\mu\)g/L\tabularnewline
WHO & Recreational: High Prob. of Effect & 20-2000
\(\mu\)g/L\tabularnewline
WHO & Recreational: Very High Prob. of Effect & \textgreater{}2000
\(\mu\)g/L\tabularnewline
\bottomrule
\end{longtable}

@@

\hypertarget{ref-usepa2009national}{}
14. USEPA (2009) National lakes assessment: A collaborative survey of
the nation's lakes. EPA 841-R-09-001.

\hypertarget{ref-nlaux5ffieldops}{}
15. USEPA (2007) Survey of the nation's lakes field operations manual
EPA841-B-07-004.

\hypertarget{ref-paul2011probability}{}
16. Paul JF, Munns WR (2011) Probability surveys, conditional
probability, and ecological risk assessment. Environmental Toxicology
and Chemistry 30: 1488--1495.

\hypertarget{ref-paul2005development}{}
17. Paul JF, McDonald ME (2005) Development of empirical, geographically
specific water quality criteria: A conditional probability analysis
approach. Journal of the American Water Resources Association 41:
1211--1223.

\hypertarget{ref-hollister2008cprob}{}
18. Hollister JW, Walker HA, Paul JF (2008) CProb: A computational tool
for conducting conditional probability analysis. Journal of
Environmental Quality 37: 2392--2396.

\hypertarget{ref-miller2013spatiotemporal}{}
19. Miller TR, Beversdorf L, Chaston SD, McMahon KD (2013)
Spatiotemporal molecular analysis of cyanobacteria blooms reveals
microcystis-aphanizomenon interactions. PloS one 8: e74933.

\hypertarget{ref-marion2012vivo}{}
20. Marion JW, Lee J, Wilkins III J, Lemeshow S, Lee C, et al. (2012) In
vivo phycocyanin flourometry as a potential rapid screening tool for
predicting elevated microcystin concentrations at eutrophic lakes.
Environmental science \& technology 46: 4523--4531.

\hypertarget{ref-beaver2014land}{}
21. Beaver JR, Manis EE, Loftin KA, Graham JL, Pollard AI, et al. (2014)
Land use patterns, ecoregion, and microcystin relationships in uS lakes
and reservoirs: A preliminary evaluation. Harmful Algae 36: 57--62.

\hypertarget{ref-wagner2011landscape}{}
22. Wagner T, Soranno PA, Webster KE, Cheruvelil KS (2011) Landscape
drivers of regional variation in the relationship between total
phosphorus and chlorophyll in lakes. Freshwater Biology 56: 1811--1824.

>>>>

Over the last decade, numerous events and legislative activities have
raised the public awareness of harmful algal blooms {[}1--3{]}. In
response the US Environmental Protection Agency (USEPA) has recently
released suggested microcystin (one of the more common toxins)
concentrations that would trigger health advisories {[}4--6{]}.
Additionally, the World Health Organization (WHO) has microcystin
advisory levels for drinking water and for a range of recreational risk
levels {[}7,8{]}. While these levels and associated advisories are
likely to help mitigate the impacts from harmful algal blooms, they are
not without complications.

@@

Fortunately, cyanobacteria and microcystin-LR has been shown to be
associated with several other, more commonly measured and well
understood components of water quality that are readily assessed in the
field {[}11{]}. For instance, there are small or hand held fluorometers
that measure chlorohpyll \emph{a}. Additionally, chlorophyll \emph{a} is
a very commonly measured component of water quality that is also known
to be positively associated with microsystin-LR concentrations
{[}12,13{]}. Recently, Yuan et. al {[}13{]} explored these associations
in detail and controlled for other related variables. In their analysis
they find that total nitrogen and chlorophyll \emph{a} show the
strongest association with microcystin. Furthermore, they identify
chlorophyll \emph{a} and total nitrogen concentrations that are
associated with exceeding 1 \(\mu\)g/L of microcystin. These findings
suggest that chlorophyll \emph{a} concentrations could also track the
new USEPA microcystin health advisory levels for drinking water.
Identifying this association would provide an important tool for water
resource managers to help manage the threat to public health posed by
cHABs and would be especially useful in the absence of measured
microcystin concentrations.

In fact, this is a similar tact to the World Health Organization who, in
addition to advisory levels for microcystin, have also proposed related
advisory levels for cyanobacteria abundance and chlorophyll \emph{a}
{[}7,8{]}. The chlorophyll \emph{a} concentrations proposed by the WHO
are for low (\textless{} 10 \(\mu\)g/L), moderate (between 10 and 50
\(\mu\)g/L, high (between 50 and 5000 \(\mu\)g/L), and very high risk
(\textgreater{}5000 \(\mu\)g/L) {[}8{]}. While these advisories have
proven to be useful tools they do suffer from being coarse, broad, and
have been found to overestimate actual risk {[}14{]}.

In this paper we build on these past efforts and utilize the National
Lakes Assessment (NLA) data and identify chlorophyll \emph{a}
concentrations that are associated with higher probabilities of
exceeding several microcystin health advisory concentrations
{[}6,8,15{]}. We build on past studies by exploring associations with
the newly announced advisory levels and by also applying a different
method, conditional probability analysis. Utilizing different methods
strengthens the evidence for suggested chlorophyll \emph{a} levels that
are associated with increased risk of exceeding the health advisory
levels as those levels are not predicated on a single analytical method.
So that others may repeat or adjust this analysis, the data, code, and
this manuscript are freely available via
\href{https://github.com/USAPE/microcystinchla}{https://github.com/USEPA/microcystinchla}.

@@

We used the 2007 NLA chlorophyll \emph{a} and microcystin-LR
concentration data {[}15{]}. These data represent a snapshot of water
quality from the summer of 2007 for the conterminous United States and
were collected as part of an ongoing probabilistic monitoring program
{[}15{]}. Water quality data, including chlorophyll \emph{a} and
microcystin-LR were obtained via an intergated sample taken from the
surface of the lake down to 2 meters. Samples were taken at the same
time from the index site (e.g.~near the centroid of the lake) and these
provide the source for both chlorophyll \emph{a} and microcystin-LR
{[}15{]}.

@@

We used a conditional probability analysis (CPA) approach to explore
associations between chlorophyll \emph{a} concentrations and World
Health Organization (WHO) and USEPA microcystin health advisory levels
{[}17{]}. Many health advisory levels have been suggested (Table
\ref{tab:microcystin_levels}), but lakes with higher microcystin-LR
concentrations in the NLA were rare. Only 1.16\% of lakes sampled had a
concentration greater than 10 \(\mu\)g/L. Thus, for this analysis we
focused on the microcystin concentrations that are better represented in
the NLA data. These were the USEPA children's (i.e.~bottle fed infants
to pre-school age children) drinking water advisory level of 0.3
\(\mu\)g/L (USEPA Child), the WHO drinking water advisory level of 1
\(\mu\)g/L (WHO Drinking), the USEPA adult (i.e.~beyond pre-school aged
individuals) drinking water advisory level of 1.6 \(\mu\)g/L (USEPA
Adult), and the WHO recreational, low probability of effect advisory
level of 2 \(\mu\)g/L (WHO Recreational).

Conditional probability analysis provides information about the
probability of observing one event given another event has also occured.
For this analysis, we used CPA to examine how the conditional
probability of exceeding one of the health advisories changes as
chlorophyll \emph{a} increases in a lake. We expect to find higher
chlorohpyll \emph{a} concentrations to be associated with higher
probabilities of exceeding the microcystin health advisory levels. We
also calculated bootstrapped 95\% confidence intervals (CI) using 10,000
bootstrapped samples. Thus, to identify chlorophyll \emph{a}
concentrations of concern we identified the value of the upper 95\% CI
across a range of conditional probabilities of exceeding each health
advisory level. Using the upper confidence limit to identify a threshold
is justified as it ensures that a given threshold is unlikely to miss a
microcystin exceedance.

As both microcystin-LR and chlorophyll \emph{a} values were highly right
skewed, a log base 10 transformation was used. Additional details of the
specific implementation are available at
\url{https://github.com/USEPA/microcystinchla}. A more detailed
discussion of CPA is beyond the scope of this paper, but see Paul et al.
{[}18{]} and Hollister et al. {[}19{]} for greater detail. All analyses
were conducted using R version 3.2.3 and code and data from this
analysis are freely available as an R package at
\url{https://github.com/USEPA/microcystinchla}.

@@

In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225
\(\mu\)g/L. Microcystin-LR concentrations of 0.05 \(\mu\)g/L represent
the detection limits. Any value greater than that indicates the presence
of microcystin-LR. Of those lakes with microcystin-LR, the median
concentration was 0.51 \(\mu\)g/L and the mean was 3.17 \(\mu\)g/L. Of
all lakes sampled, 21\% of lakes exceeded the USEPA Child level, 8.8\%
of lakes exceeded the USEPA Adult level, 11.7\% of lakes exceeded the
WHO Drinking level, and 7.3\% of lakes exceeded the WHO Recreational
level. Chlorophyll \emph{a}, ranged from 0.07 to 936 \(\mu\)g/L and this
captures the range of trrophic states from oligotrophic to
hypereutrophic. All lakes had detectable levels of chlorophyll \emph{a}.
The median concentration was 7.79 \(\mu\)g/L and the mean was 29.63
\(\mu\)g/L. The association between chlorophyll \emph{a} and the upper
confidence interval across a range of conditional probability values are
shown in Table \ref{tab:mc_chla_table}. Specific chlorophyll \emph{a}
concentrations that are associated with greater than even odds of
exceeding the advisory levels were 23, 68, 84, and 104 \(\mu\)g/L for
0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory levels, respectively (Table
\ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).

The chlorophyll \emph{a} cutoffs may be used to predict whether or not a
lake exceeds the microcystin health advisories. Doing so allows us to
compare the accuracy of the prediction as well as evaluate false
negatives. Total accuracy of the four cutoffs predicting microcystin
exceedances were 74\% for the USEPA children's drinking water advisory,
86\% for the WHO drinking water advisory, 89\% for the USEPA adult
drinking water advisory, and 91\% for the WHO recreational advisory
(Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}).
However, total accuracy is only one part of the prediction performace
with which we are concerned.

@@

A second potential use is to identify past bloom events from historical
data. As harmful algal blooms are made up of many species and have
various mechanisms responsible for adverse impacts (e.g., toxins,
hypoxia, odors), there is no single definition of a bloom. For cHABs,
one approach has been to utilize phycocyanin to screen for or identify
bloom events {[}20{]}. This is a useful approach, but phycocyanin is not
always available, thus limiting its utility especially for examining
historical data. Using our chlorophyll \emph{a} cutoffs provides a value
that is also associated with microcystin-LR and can be used to classify
lakes, from past surveys, as having bloomed.

The values we propose are national and may miss regional variation in
water quality, including, chlorophyll \emph{a} and microcystin-LR
{[}22{]}. A set of regional conditional probabilities would be
interesting; however, limiting the analysis to the data available per
region would make interpretation difficult. The sample size for each of
the regional conditional probabilities would be reduced and the number
of lakes in each region that exceed the microcystin values would also be
reduced. Thus, our confidence in the conditional probabilites would be
less (i.e.~greatly increased confidence intervals) and the relationships
less pronounced as we have fewer lakes on which to base the
probabilites. Thus, this dataset is best for making national scale
recommendations.

There are two other limitations with the 2007 National Lakes Assessment
dataset. First, it represents a single sample from a lake and does not
capture temporal dynamics. Second, validation of the predictions with
the 2007 data alone would be challenging as the data would need to be
subset and this would only sever to increase the uncertainty of our
conditional probabilities, reducing our ability to validate the presence
of microcystin-LR. The 2017 National Lakes Assessment would be ideal for
this task. However, as of this writing, the 2012 National Lakes
Assessment data are not public. When these data are released, a
validation of this approach can be completed then.

Lastly, using chlorophyll \emph{a} is not meant as a replacement for
testing of microcystin-LR or other toxins. It should be used when other,
direct measurements of cyanotoxins are not available. In those cases,
which are likely to be common at least in the near future, using a more
ubiquitous measurement such as chlorophyll \emph{a} will provide a
reasonable proxy for the probability of exceeding a microcystin health
advisory level and provide better protection against adverse effects in
both drinking and recreational use cases.

@@

@@

\begin{longtable}[c]{@{}lll@{}}
\caption{Various microcystin health advisory concentrations.
\label{tab:microcystin_levels}}\tabularnewline
\toprule
Source & Type & Concentration\tabularnewline
\midrule
\endfirsthead
\toprule
Source & Type & Concentration\tabularnewline
\midrule
\endhead
USEPA & Child Drinking Water Advisory & 0.3 \(\mu\)g/L\tabularnewline
WHO & Drinking Water & 1 \(\mu\)g/L\tabularnewline
USEPA & Adult Drinking Water Advisory & 1.6 \(\mu\)g/L\tabularnewline
WHO & Recreational: Low Prob. of Effect & 2-4 \(\mu\)g/L\tabularnewline
WHO & Recreational: Moderate Prob. of Effect & 10-20
\(\mu\)g/L\tabularnewline
WHO & Recreational: High Prob. of Effect & 20-2000
\(\mu\)g/L\tabularnewline
WHO & Recreational: Very High Prob. of Effect & \textgreater{}2000
\(\mu\)g/L\tabularnewline
\bottomrule
\end{longtable}

@@

\hypertarget{ref-loftin2016cyanotoxins}{}
14. Loftin KA, Graham JL, Hilborn ED, Lehmann SC, Meyer MT, et al.
(2016) Cyanotoxins in inland lakes of the united states: Occurrence and
potential recreational health risks in the ePA national lakes assessment
2007. Harmful Algae 56: 77--90.

\hypertarget{ref-usepa2009national}{}
15. USEPA (2009) National lakes assessment: A collaborative survey of
the nation's lakes. EPA 841-R-09-001.

\hypertarget{ref-nlaux5ffieldops}{}
16. USEPA (2007) Survey of the nation's lakes field operations manual
EPA841-B-07-004.

\hypertarget{ref-paul2011probability}{}
17. Paul JF, Munns WR (2011) Probability surveys, conditional
probability, and ecological risk assessment. Environmental Toxicology
and Chemistry 30: 1488--1495.

\hypertarget{ref-paul2005development}{}
18. Paul JF, McDonald ME (2005) Development of empirical, geographically
specific water quality criteria: A conditional probability analysis
approach. Journal of the American Water Resources Association 41:
1211--1223.

\hypertarget{ref-hollister2008cprob}{}
19. Hollister JW, Walker HA, Paul JF (2008) CProb: A computational tool
for conducting conditional probability analysis. Journal of
Environmental Quality 37: 2392--2396.

\hypertarget{ref-miller2013spatiotemporal}{}
20. Miller TR, Beversdorf L, Chaston SD, McMahon KD (2013)
Spatiotemporal molecular analysis of cyanobacteria blooms reveals
microcystis-aphanizomenon interactions. PloS one 8: e74933.

\hypertarget{ref-marion2012vivo}{}
21. Marion JW, Lee J, Wilkins III J, Lemeshow S, Lee C, et al. (2012) In
vivo phycocyanin flourometry as a potential rapid screening tool for
predicting elevated microcystin concentrations at eutrophic lakes.
Environmental science \& technology 46: 4523--4531.

\hypertarget{ref-beaver2014land}{}
22. Beaver JR, Manis EE, Loftin KA, Graham JL, Pollard AI, et al. (2014)
Land use patterns, ecoregion, and microcystin relationships in uS lakes
and reservoirs: A preliminary evaluation. Harmful Algae 36: 57--62.

\hypertarget{ref-wagner2011landscape}{}
23. Wagner T, Soranno PA, Webster KE, Cheruvelil KS (2011) Landscape
drivers of regional variation in the relationship between total
phosphorus and chlorophyll in lakes. Freshwater Biology 56: 1811--1824.
